---
title: "Mouse MSigDB gene sets"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: TBA

## Intro

Mouse omics experiments are nearly as common as those using human, yet
MSigDB only published human gene sets.

The msigdbr R package provides a convenient way to get these gene sets,
yet it requires a lot of R skills to get these into a form that works 
with the GSEA java app.

Previously I published a bash script to convert the GMT fil from human
to mouse, but this approach has a downside that the gene sets and gene
ortholog mapping files become outdated.

A good solution would be to use msigdbr package to generate the gmt files
for folks to download. 
This should be conducted periodically to automatically update ortholog maps
and msigdb versions.

```{r,libs}

library("msigdbr")
library("dplyr")
library("parallel")

writeGMT <- function (object, fname ){
  if (class(object) != "list") stop("object should be of class 'list'")
  if(file.exists(fname)) unlink(fname)
  for (iElement in 1:length(object)){
    write.table(t(c(make.names(rep(names(object)[iElement],2)),object[[iElement]])),
                sep="\t",quote=FALSE,
                file=fname,append=TRUE,col.names=FALSE,row.names=FALSE)
  }
}

#writeGMT(object=athl,fname="Ath_AGI_LOCUS_TAIR10_Aug2012.txt.gmt")

DATE=Sys.Date()

```

## Fetch MSigDB gene sets

```{r,fetch1}

as.data.frame(msigdbr_collections())

# GO
go_gene_sets <- msigdbr(species = "mouse", category = "C5")
gs_names <- unique(go_gene_sets$gs_name)
go_gene_sets_list <- mclapply(gs_names, function (gsn) {
  pull(unique(go_gene_sets[which(go_gene_sets$gs_name == gsn),"gene_symbol"])) },
mc.cores=16)
names(go_gene_sets_list) <- gs_names
FILENAME=paste("mouse_msigdb_go_",DATE,".gmt",sep="")
writeGMT(object=go_gene_sets_list,fname=FILENAME)

# KEGG
c2_gene_sets <- msigdbr(species = "mouse", category = "C2")
unique(c2_gene_sets$gs_subcat)
kegg_gene_sets <-  c2_gene_sets[which(c2_gene_sets$gs_subcat == "CP:KEGG"),]
gs_names <- unique(kegg_gene_sets$gs_name)
kegg_gene_sets_list <- mclapply(gs_names, function (gsn) {
  pull(unique(kegg_gene_sets[which(kegg_gene_sets$gs_name == gsn),"gene_symbol"])) },
mc.cores=16)
names(kegg_gene_sets_list) <- gs_names
FILENAME=paste("mouse_msigdb_kegg_",DATE,".gmt",sep="")
writeGMT(object=kegg_gene_sets_list,fname=FILENAME)

## Reactome
reactome_gene_sets <-  c2_gene_sets[which(c2_gene_sets$gs_subcat == "CP:REACTOME"),]
gs_names <- unique(reactome_gene_sets$gs_name)
reactome_gene_sets_list <- mclapply(gs_names, function (gsn) {
  pull(unique(reactome_gene_sets[which(reactome_gene_sets$gs_name == gsn),"gene_symbol"])) },
mc.cores=16)
names(reactome_gene_sets_list) <- gs_names
FILENAME=paste("mouse_msigdb_reactome_",DATE,".gmt",sep="")
writeGMT(object=reactome_gene_sets_list,fname=FILENAME)

## All msigDB
msigdb_gene_sets <-  msigdbr(species = "mouse")

gs_names <- unique(msigdb_gene_sets$gs_name)

msigdb_gene_sets_list <- mclapply(gs_names, function (gsn) {
  pull(unique(msigdb_gene_sets[which(msigdb_gene_sets$gs_name == gsn),"gene_symbol"])) },
mc.cores=16)
names(msigdb_gene_sets_list) <- gs_names
FILENAME=paste("mouse_msigdb_all_",DATE,".gmt",sep="")
writeGMT(object=msigdb_gene_sets_list,fname=FILENAME)

```

## Session information

```{r,sessioninfo}

sessionInfo()

```
